---
title: "R Notebook"
output: html_notebook
---

Joining multiple data. tables
A Join describe the action of combining information from two different data tables into a single data table.
The majority of R's functions for analyzing and visualizing data are designed to work on a single data frame or data table.

The first skill we need to learn is to identify the join KEY COLUMNS.
Columns that link information across two tables to use as join keys.  
install.packages(")
```{r}
demographics <- data.table(name = c("Trey", "Matthe", "Angela"),
                            gender = c(NA, "M", "F"),
                            age = c(54, 43, 39))

shipping <- data.table(name = c("Matthew", "Trey", "Angela"), 
                            address = c("7 Mill road", "12 High street", "33 Pacific boulevard"))
```

To identify join keys, we need to learn about the contents of the data tables.
Inspecting data.tables in your R sessions

- The tables() function will show all data.tables loaded in R session.
- The str() will show the type of each column in a single data.table
- Typing the variable name (data.table) will show the values stored in a data table.

=========================================================================================================================

Exploring data.tables

```{r}
# What data.tables are in my R session?
tables()

# View the first six rows 
head(netflix)
head(imdb)

# Print the structure
str(netflix)
str(imdb)
```

#=================================================== MERGE FUNCTION =====================================================

All four standard joins taht originally come from database query languages, such as SQL. We will learn how to do these
joins using the merge() function. This function comes from base R but is extended ind the data table package.

- Inner Join: keep observations that have information in BOTH data.tables
An inner join is the DEFAULT behavior of the merge() function.
```{r}
merge(x = demographics, y = shipping,  by.x = "name", by.y="name")
```
demographics: name, gender, age
shipping:     name, address
output:       name, gender, age, address

- Full Join |Full outer join: keep all observations that are in either data.table
setting the all argument to be equal to TRUE.

```{r}
merge(x = demographics, y = shipping, by = "name", all = TRUE)
```

#=================================================== INNER JOIN  =======================================================
```{r}
# Inner join netflix and imdb
merge(netflix, imdb, by = "title")
```
> # Inner join netflix and imdb
> merge(netflix, imdb, by = "title")
                            title episodes release_date rating
1:                 13 Reasons Why       13    31-Mar-17    8.4
2: A Series of Unfortunate Events        8    13-Jan-17    7.9
3:                          Gypsy       10    30-Jun-17    7.0
4:                     Mindhunter       10    13-Oct-17    8.9
5:                          Ozark       10    21-Jul-17    8.5
#=================================================== FULL JOIN  =======================================================

```{r}
# Full join netflix and imdb
merge(netflix, imdb, by = "title", all = TRUE)
```

#================================================= LEFT AND RIGHT JOINS ================================================

A left join keeps only the observation that are present in the data table on the left side of the join

- Left join: all.x = TRUE
```{r}
merge(x = demographics, y = shipping, by ="name", all.x = TRUE)
```

- Reight join: all.y = TRUE
```{r}
merge(x = demographics, y = shipping, by = "name", all.y = TRUE)
```

Default values
  Default values for all, all.x and all.y are FALSE in the mege() function
  Look up function argument defaults using help("merge")
  
#================================================= JOINS  ==========================================================

```{r}
# Left join imdb to netflix
merge(x = netflix, y = imdb, by = "title", all.x = TRUE)
```

```{r}
# Right join imdb to netflix
merge(x = netflix, y = imdb, name = "title", all.y = TRUE)
```

```{r}
# Identify the key for joining capitals and population
capitals_population_key <- "city"

# Left join population to capitals
capital_pop <- merge(capitals, population, by = capitals_population_key, all.x = TRUE)
capital_pop
```

```{r}
# Identify the key for joining capital_pop and area
capital_pop_area_key <- "state"

# Inner join area to capital pop
australia_stats <- merge(capital_pop, area, by = capital_pop_area_key)

# Print the final result
australia_stats
```

