---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
output: html_notebook
---

Joining Data With Data.Table in R - Incorporating joins into data.table workflow

Chaining data.table expressions:
data.table expressions can be chained in sequence
demographics[...][...]

General form of chainin a join:

DataTable1[DataTable2, on][i, j, by]
           #join to which data.table
                       #join key columns
                           #on which rows?
                              #what to do?
                                  #grouped by what?

The most flexible way of incorporating joins into our DataTable workflows is by chaining DataTable expressions.
  "Join DataTable1 to DataTable2 using the column in "on" as the join key.
    Then in the result, filter by rows in "i", compute on columns "j" grouped by values in columns "by". 
  
============================================================================================================

To get the average spent per sale for return customers grouped by gender, we chain two data table expressions.

customers[purchases, on =.(name)][sales > 1, j = .(avg_spent = sum(spent) / sum(sales)), by = .(gender)]

*In the first, we join customers to purchases, then in the second, we filter to customers with more than 1, sale,
  calculate the average spent per sale in j, grouped by the gender column in by.

    gender    avg_spent
1:    M          14
2:    F          20

============================================================================================================

Computation with joins: Efficient for large data.tables

Using the j argument to compute on columns in the same DataTable expression as a join. 

DataTable1[DataTable2, on, j]
            #join to which data.table?
                      #using which columns as keys?
                           #what to do on the join result?
                        
This save a lot of time and memory for large DataTables
                          
============================================================================================================

Joining and column creation

customers[purchases, on = .(name), return_customer := sales > 1]
customers

============================================================================================================

Grouping by matches

DataTable1[DataTable2, on, j, by = .EACHI]
          #joint to which table?
                       #using which columns as keys?
                           #what to do on the join result?
                              #grouped by each match in DataTabel
                              
" Join DataTable1 to DataTable 2, then for each row in DataTable2 group calculations in j 
      by the matching rows in DataTable1

#=========================================== .EACHI========================================================    
Using by = .EACHI is a useful way of checking our join has the number of matches you expect, and finding rows 
that may cause problems in downstream analysis. Remember if we want to use by= to group computations in j by
another column we will need to chain a second data.table[] expression.
      
```{r}
shipping[customers, on = .(name), j = .("#of shipping addresses" = .N), by = .EACHI]
```

```{r}
# Inner join capitals to population
 population[capitals, on = .(city), nomatch =0]
```

```{r}
# Join and sum
population[capitals, on = .(city), nomatch = 0,
           j = sum(percentage)]
```

```{r}
# How many continents is each country listed in?
continents[life_exp, on = .(country), .N, 
           by = .EACHI]
```

```{r}
# What countries are listed in multiple continents?
continents[life_exp, on = .(country), .N, 
           by = .EACHI][N > 1]
```
World Life Expectancy
```{r}
# Calculate average life expectancy per continent:
avg_life_expectancy <- continents[life_exp, on = .(country), 
                                  nomatch = 0][, j = mean(years), 
                                               by = .(continent)]
avg_life_expectancy
```

