---
title: "R Notebook"
output: html_notebook
---

Joining Data With Data.Table in R - data.table syntax

General form of data.table syntax
DT[i, j, by]
DT[on which rows?, what to do?, grouped by what?]
"Take DT, filter its rows in i, then compute on columns j, grouped by values in columns by. 

General form of data.table joins
DT[i, on]
DT[join to which data.table, join key columns]
"Join DT to i using the column in on as the join key"

The on argument
Variables inside list() or .() are looked up in the column names of both data.tables
shipping[demographics, on = list(name)]
shipping[demographics, on = .(name)]

Left joins
```{r}
shipping[demographics, on = .(name)]
```
Right joins
```{r}
demographics[shipping, on = .(name)]
```
Inner joins
```{r}
shipping[demographics, on = .(name), nomatch = 0]
#nomatch tells the data table syntax to ignore rows that cannot be matched between both
```
Full joins
```{r}
merge(demographics, shipping, by = "name", all = TRUE)
```
Anti-joins
```{r}
demographics[!shipping, on = .(name)]
#contains only rows that have no match in the data table on the right
```
Character vectors can also be used to specify the join key
```{r}
join_key <- c("name")
shipping[demographics, on = join_key]
```

===========================================================================================

Right join with the data.table syntax
```{r}
# Right join population to capitals using data.table syntax
capitals[population, on = .(city)]
```
Right join using merga
```{r}
# Right join using merge
merge(x = capitals, y = population , by = "city", all.y = TRUE)
```
Inner join with data.table syntax
```{r}
# Inner join with the data.table syntax
capitals[population, on = .(city), nomatch = 0]
```
Anti-join
```{r}
# Anti-join capitals to population
population[!capitals, on = .(city)]
```
```{r}
# Anti-join capitals to area
area[!capitals, on = .(state)]
```

