---
title: "R Notebook"
output: html_notebook
---

Joining Data with Data.Table in R - Duplicate Matches

The lest set of join problems we will commonly encounter arise from DUPLICATE MATCHES  in the join key values.

Sometimes we will want to join two data tables based on columns with DUPLICATE ENTRIES, leading to multiplicative matches.

Some error is shown to protect us from making unexpectedly large data tables by accident because MULTIPLICATIVE MATCHES are a common problem in MISSPECIFIED joins. 

=================================================================

Allowing multiplicativa matches - allow.cartesian

This can set allow.cartesian to proceed intentionally. 


=================================================================

Missing Values (na) will match all other missing values

If we have missing values in our join column, these can also lead to multiplicative matches, because they will match all other missing values in the join. 

! To remove these mathce, which are likely MEANINGLESS, we can FILTER OUT rows with missing values in the join key columns. 

Keeping only the first match - The Data Table we only want to match the most recent collection. To do this we can set mult = "first" to keep only the first matching row in the main DT. 
```{r}
site1[site2, on = .(songs), mult = "first"]
```

=================================================================

Identifying and removing duplicates

duplicated(): what rows are duplicates?

! The duplicated() function from the data table package wil tell WHICH ROWS ARE DUPLICATES. It returns a vector of TRUE and FALSE, values, where each entry corresponds to a row in the data table.

A value of TRUE means that the rows has the SAME VALUES AS ONE OF THE PREVIOUS ROWS in the data table. 

An additional argument, "by", can be used to restric checks to only a SUBSET of columns, for example looking for duplicates in just the join key(CPF)column.  


================================================================

The unique() function

unique(): filter data.table to just unique rows

The unifque fucntion takes the same arguments, but instead returns the data table without the rows that were determined to be duplicates.

================================================================

Joining with missing values

• Using the merge() function, inner join cardio to heart with the appropriate argument to override any errors that you encounter.

```{r}
# Try an inner join
merge(heart, cardio, by = "gene", allow.cartesian = TRUE)
```

• Filter MISSING values
```{r}
heart_2 <- heart[!is.na(gene)]
cardio_2 <- cardio[!is.na(gene)]
```

• INNER JOIN the filtered data.tables
```{r}
merge(heart_2, cardio_2, by = "gene")
```

Filtering Duplicates

• Use the unique() function to remove duplicate entries in the "gene" column in both heart_2 and cardio_2. Keep only the last row for each gene.

```{r}
heart_3 <- unique(heart_2, by = "gene", fromLast = TRUE)
cardio_3 <- unique(cardio_2, by = "gene", fromLast = TRUE)
```

• Inner join cardio_3 to heart_3 using the merge() function. Append ".heart" and ".cardio" as suffixes to the "change" and "pvalue" columns.

```{r}
# Inner join
reproducible <- merge(heart_3, cardio_3, by = "gene", suffixes = c(".heart", ".cardio"))
reproducible
```

================================================================

Joining and Filtering Duplicates

• Using the data.table syntax, right join framingham to heart_2, taking the first probe for each gene in heart_2.

```{r}
# Right join taking the first match
heart_2[framingham, on = .(gene), mult = "first"]
```

• Anti-join framingham to reproducible to see whether you found any genes that have not previously been seen to have an association with heart disease.

```{r}
# Anti-join
reproducible[!framingham, on = .(gene)]
```

