---
title: "R Notebook"
output: html_notebook
---

Joining Data with Data.Table in R - Melting Data.Tables

In this lesson, we will learn hot to melt data tables from wIDE TO LONG formats.



MELT() FUNCTION
====================================================================

melt() to reshape a WIDE DATA TABLE to a LONG data table. COLUNA P/ LINHAS.

```{r}
melt(sales_wide, measure.vars = c("2015","2016"))
```
   quarter variable value
1:       1     2015 3200100
2:       2     2015 2950000
3:       3     2015 2980700
4:       4     2015 3420000
5:       1     2016 3350000
6:       2     2016 3000300
7:       3     2016 3120200
8:       4     2016 3670000

melt() will take these columsn and stack them on top of each other to new columns: "variable" and "value".

```{r}
melt(sales_wide, id.vars = "quarter", 
     variable.name = "year", value.name = "amount")
```

   quarter year  amount
1:       1 2015 3200100
2:       2 2015 2950000
3:       3 2015 2980700
4:       4 2015 3420000
5:       1 2016 3350000
6:       2 2016 3000300
7:       3 2016 3120200
8:       4 2016 3670000

=====================================================================

Melting a wide table

Use the melt() function to create a long format data.table stacking the values for each country's column.

```{r}
# Reshape gdp_per_capita to the long format
melt(gdp_per_capita, id.vars = "year")
```

Within your call to melt(), name the new "variable" column as "country" and the new "value" column as "gdp_pc".

```{r}
# Rename the new columns
melt(gdp_per_capita, id.vars = "year", 
     variable.name ="country", value.name = "gdp_pc")
```

Use the measure.vars argument to stack the columns "Week_50" and "Week_51". The new column containing its values should be called "cases", and the new column of variable labels should be called "period".

```{r}
# Stack Week_50 and Week_51
melt(ebola_wide, measure.vars = c("Week_50", "Week_51"), 
     variable.name = "period", value.name = "cases")
```

Modify the call to melt() to also ask to keep only the "Location" column as row identifiers, dropping the "period_start" and "period_end" columns from the result.

```{r}
# Modify the code
melt(ebola_wide, measure.vars = c("Week_50", "Week_51"), 
     variable.name = "period", value.name = "cases", 
     id.vars = "Location")
```
     Location  period cases
 1:       BEYLA Week_50    NA
 2:     CONAKRY Week_50    37
 3:     CONAKRY Week_50    NA
 4:       COYAH Week_50    31
 5:       COYAH Week_50    NA....
