---
title: "R Notebook"
output: html_notebook
---

Joining Data with Data.Table in R - Casting data.tables

In this lesson, we will learn hot to cast data tables from LONG TO WIDE formats. 

====================================================================

Long data.table

```{r}
sales_wide <- dcast(sales_long, quarter ~ year, value.var = "amount")
```

dcast() function allows to make the reverse transformation, reshaping a long format data table into a wide one. 

dcast() takes 3 arguments
  1. data.table to reshape
  2. rows to keep behind as identifiers
  3. group labels to split by

====================================================================

  Use dcast() to build a wide data.table containing the population of each country, with each country as a row and each year as a column.

```{r}
# Split the population column by year
dcast(gdp_oceania, formula = country ~ year, value.var = "population")
```

Create another wide format data.table containing the GDP of each country with each year as a row and each country as a column.

```{r}
# Split the gdp column by country
dcast(gdp_oceania, formula = year ~ country, value.var = "gdp")
```

Convert gdp_oceania to wide format with each country as a row, and splitting both the "gdp" and "population" columns into columns for each "year".

```{r}
# Split the gdp and population columns by year
dcast(gdp_oceania, formula = country ~ year, value.var = c("gdp", "population"))
```

Convert wide to a matrix using the "country" column as the matrix rownames.

```{r}
# Reshape from wide to long format
wide <- dcast(gdp_oceania, formula = country ~ year, value.var = c("gdp", "population"))

# convert to a matrix
as.matrix(wide, rownames = "country")
```

Modify the call to dcast() so that the "continent" column is also kept aside as row identifiers.

```{r}
# Modify your previous code
dcast(gdp_oceania, formula = country + continent ~ year, value.var = c("gdp", "population"))
```

Convert gdp_by_industry_oceania to a wide format with each country as a row, splitting the "gdp" column so that the result has columns containing the gdp for each industry and each year.


```{r}
# Split gdp by industry and year
dcast(gdp_by_industry_oceania, country ~ industry + year, value.var = "gdp")
```
